Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-14 write-ups/PLDI/paper.tex
Shortenings.

\paragraph{The needs of irregular data-parallel algorithms}
%Acceleration of general-purpose computations on graphics processing
%units (GPUs) has tended to focus on \emph{regular} data-parallel
%algorithms, for which work to be processed can be evenly split between
%the workgroups that execute a GPU kernel ahead of time.  However,
Many
interesting data-parallel algorithms are \emph{irregular}: the amount
of work to be processed is unknown ahead of time and may be determined
by the computation itself.  The last decade has seen growing interest in accelerating irregular
data-parallel algorithms on GPUs, particularly algorithms that
traverse and manipulate linked data structures~\cite{owens-persistent,TPO10,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/ipps/DavidsonBGO14,DBLP:conf/hipc/HarishN07,DBLP:journals/topc/MerrillGG15,DBLP:conf/egh/VineetHPN09,DBLP:conf/ppopp/NobariCKB12,DBLP:conf/hpcc/SolomonTT10a,DBLP:conf/popl/PrabhuRMH11,DBLP:conf/ppopp/Mendez-LojoBP12,DBLP:conf/oopsla/PaiP16,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/oopsla/SorensenDBGR16}.  \ADComment{Tyler, can you add relevant workstealing papers if not here already?}

@@

\paragraph{Work stealing example}
%
The work stealing paradigm is useful when a computation can be split
into multiple tasks that can be processed independently, but where the
number of tasks is not known ahead of time because processing a given
task may create an arbitrary number of new tasks. Work stealing
enables dynamic balancing of tasks across several processing units,
OpenCL workgroups in our case. Each workgroup has an associated task
queue from which it obtains tasks to process, and to which it stores
newly created tasks. When a workgroup's queue becomes empty, the
workgroup---rather than terminating---tries to \emph{steal} a task
from an other queue. Workgroups keep on processing tasks until all the
work is done. Previous work addresses the feasibility and efficiency
of work stealing on GPUs~ref\cite{dlb-web,TPO10}.

\myfiglong\ref{fig:workstealing} shows a simplified version of an OpenCL
kernel that uses a work stealing queue. All threads start their
execution at the function marked with the $\mathsf{kernel}$ keyword
(line~\ref{line:wksteal:kernelfunc}). They receive a pointer to the task
queues, which reside in global memory, and which were initialized by the
host program beforehand to contain the initial tasks. Each thread can
identify the workgroup it belongs to via the function $\getgroupid()$
(line~\ref{line:wksteal:getgroupid}). Here the group id is stored into
the variable $\mathsf{queue\_id}$ (which is, by default, in private
memory) and used to access the relevant task queue. The
$\mathsf{pop\_or\_steal}$ function (line~\ref{line:wksteal:poporsteal})
is called by all threads of a workgroup.  Inside
$\mathsf{pop\_or\_steal}$ only the thread with a local id equal to $0$,
also called the \emph{master thread} of the workgroup, will try to pop a
task from the workgroup's queue, or steal a task from a different
queue. The master thread uses local memory inside
$\mathsf{pop\_or\_steal}$ to ensure that the returned value is valid for
all threads of its workgroup. If a task was obtained, then all threads
of the workgroup participate in the processing of the task via the
$\mathsf{process\_task}$ function (line~\ref{line:wksteal:processtask}).
Processing this task may lead to the creation of further tasks, which
the master thread will push to the workgroup's queue.

>>>>

\paragraph{The needs of irregular data-parallel algorithms}
%Acceleration of general-purpose computations on graphics processing
%units (GPUs) has tended to focus on \emph{regular} data-parallel
%algorithms, for which work to be processed can be evenly split between
%the workgroups that execute a GPU kernel ahead of time.  However,
Many
interesting data-parallel algorithms are \emph{irregular}: the amount
of work to be processed is unknown ahead of time and may be determined
by the computation itself.  The last decade has seen growing interest in accelerating irregular
data-parallel algorithms on GPUs, particularly algorithms that
traverse and manipulate linked data structures~\cite{owens-persistent,TPO10,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/ipps/DavidsonBGO14,DBLP:conf/hipc/HarishN07,DBLP:journals/topc/MerrillGG15,DBLP:conf/egh/VineetHPN09,DBLP:conf/ppopp/NobariCKB12,DBLP:conf/hpcc/SolomonTT10a,DBLP:conf/popl/PrabhuRMH11,DBLP:conf/ppopp/Mendez-LojoBP12,DBLP:conf/oopsla/PaiP16,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/oopsla/SorensenDBGR16,dlb-web,TPO10}.

@@

\paragraph{Work stealing example}
%
Work stealing enables dynamic balancing of tasks across several
processing units, and is useful when the number of tasks to be
processed in parallel is dynamic, because processing
one task may create an arbitrary number of new tasks.  In the context
of GPUs, each workgroup has an associated queue from which it
obtains tasks to process, and to which it stores newly created
tasks. If a workgroup's queue is empty, the workgroup tries to \emph{steal} a task from an other
queue. Workgroups keep on processing tasks until all the work is
done. Previous work addresses the feasibility and efficiency of work
stealing on GPUs~\cite{dlb-web,TPO10}.

\myfiglong\ref{fig:workstealing} shows a simplified version of an OpenCL
kernel that uses a work stealing queue. All threads start their
execution at the function marked with the $\mathsf{kernel}$ keyword
(line~\ref{line:wksteal:kernelfunc}). They receive a pointer to the task
queues, in global memory, initialized by the
host to contain the initial tasks. Each thread identifies
the workgroup it belongs to via the function $\getgroupid()$
(line~\ref{line:wksteal:getgroupid}). Here the group id is stored into
the variable $\mathsf{queue\_id}$ (which is, by default, in private
memory) and used to access the relevant task queue. The
$\mathsf{pop\_or\_steal}$ function (line~\ref{line:wksteal:poporsteal})
is called by all threads of a workgroup.  Inside
$\mathsf{pop\_or\_steal}$ only the thread with a local id equal to $0$,
the \emph{master thread}, will try to pop a
task from the workgroup's queue, or steal a task from a different
queue. The master thread uses local memory inside
$\mathsf{pop\_or\_steal}$ to ensure that the returned value is valid for
all threads of its workgroup. If a task was obtained, then all threads
of the workgroup participate in the processing of the task via 
$\mathsf{process\_task}$ (line~\ref{line:wksteal:processtask}).
Processing this task may lead to the creation of further tasks, which
the master thread will push to the workgroup's queue.
